---
title: "Cognition Measures in the British Birth Cohorts"
author: "Liam Wright"
date: "16 November 2021"
output: html_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
```

# Introduction

In this document, I show code to clean individual cognition measures in the British Birth Cohorts. For the NCDS, BCS70 and MCS, the raw data files were downloaded from the UK Data Service (End User Licence versions). Data for the NSHD were obtained from David Bann. I clean cognition variables that were given to all sample members for a given cohort, rather than those given to sub-samples or to the children of sample members.

The source information for the cognition measures was drawn from Vanessa Moulton's ([2020](https://www.closer.ac.uk/cross-study-data-guides/cognitive-measures-guide/)) guide to cognitive measures in five British birth cohort studies written for CLOSER.

```{r}
library(tidyverse)
library(haven)
library(labelled)
library(skimr)
library(scales)
library(glue)

negative_to_na <- function(x){
  ifelse(x < 0, NA, x)
}

get_correct <- function(df, vars, new_var, 
                        correct_func = function(x) ifelse(x == 1, 1, 0)){
  df %>%
    select(id = 1, {{ vars }}) %>%
    mutate(across(-id, 
                  list(correct = correct_func,
                       observed = ~ ifelse(is.na(.x) | .x < 0, 0, 1)))) %>%
    select(id, matches("(correct|observed)")) %>%
    rowwise() %>%
    summarise(id = id, 
              correct = sum(c_across(matches("_correct")), na.rm = TRUE),
              observed = sum(c_across(matches("_observed")), na.rm = TRUE),
              .groups = "drop") %>%
    mutate(correct = ifelse(observed != 0, correct, NA)) %>%
    select(id, {{new_var}} := correct)
}

get_descriptives <- function(df){
  my_skim <- skim_with(numeric = sfl(n = ~ sum(!is.na(.)),
                                     missing = ~ sum(is.na(.)),
                                     mean = ~ mean(., na.rm = TRUE),
                                     sd = ~ sd(., na.rm = TRUE),
                                     median = ~ median(., na.rm = TRUE),
                                     min = ~ min(., na.rm = TRUE),
                                     max = ~ max(., na.rm = TRUE)),
                       base = sfl(),
                       append = FALSE)
  my_skim(df, -1) %>%
    yank("numeric")
}

plot_cog <- function(df){
  df %>% 
    pivot_longer(-1) %>%
    ggplot() + 
    aes(x = value) +
    facet_wrap(~ name, scales = "free") +
    scale_y_continuous(labels = comma) +
    geom_bar() +
    labs(x = NULL, y = NULL)
}

plot_cor <- function(df){
  df_cor <- df %>%
    rename(id = 1) %>%
    pivot_longer(-1) %>%
    full_join(., ., by = "id") %>%
    filter(name.x < name.y) %>%
    drop_na()
  
  cor_df <- df_cor %>% 
    group_by(name.x, name.y) %>% 
    summarise(cor = cor(value.x, value.y, method = "spearman"),
              .groups = "drop") %>%
    mutate(string = glue("corr = {round(cor, 2)}"))
  
  ggplot(df_cor) +
    aes(x = value.x, y = value.y) +
    facet_grid(name.y ~ name.x, scales = "free") +
    geom_jitter(alpha = 0.05, color = "grey60") +
    geom_smooth() +
    geom_text(data = cor_df, aes(label = string, x = Inf, y = Inf),
              color = "red", hjust = 1.1, vjust = 1.1) +
    labs(x = NULL, y = NULL)
}
```

# National Child Development Study (1958 cohort)

## Individual Cognition Measures
### Age 7
```{r}
ncds_0016_raw <- read_dta("D:/NCDS/0y-16y/ncds0123.dta")

ncds_07 <- ncds_0016_raw %>%
  mutate(
    reading_07 = negative_to_na(n92), # Southgate Group Reading Test; Verbal (Reading)
    arithmetic_07 = negative_to_na(n90), # Problem Arithmetic Test; Arithmetic
    visual_07 = negative_to_na(n457), # Copying Designs Test; Visual-Spatial
    drawing_07 = negative_to_na(n1840), # Human Figure Drawing; General Ability (Perceptual)
  ) %>%
  select(id = 1, matches("_07"))
```

```{r, echo = FALSE}
get_descriptives(ncds_07)
plot_cog(ncds_07)
plot_cor(ncds_07)
```

### Age 11
```{r}
ncds_11 <- ncds_0016_raw %>%
  mutate(
    verbal_11 = negative_to_na(n914), # General Ability Test; Verbal Reasoning
    nonverbal_11 = negative_to_na(n917), # General Ability Test; Verbal Reasoning
    reasoning_11 = negative_to_na(n920), # General Ability Test; Sum of Verbal and Non-Verbal
    reading_11 = negative_to_na(n923), # Reading Comprehension Test; Verbal (Reading)
    arithmetic_11 = negative_to_na(n926), # Mathematics Test; Arithmetic
    visual_11 = negative_to_na(n929), # Copying Designs Test; Visual-Spatial
  ) %>%
  select(id = 1, matches("_11"))
```

```{r, echo = FALSE}
get_descriptives(ncds_11)
plot_cog(ncds_11)
plot_cor(ncds_11)
```
Note, the excess of scores of zero in the verbal and non-verbal reasoning tests.

### Age 16
```{r}
ncds_16 <- ncds_0016_raw %>%
  mutate(
    reading_16 = negative_to_na(n2928), # Reading Comprehension Test; Verbal (Reading)
    maths_16 = negative_to_na(n2930), # Mathematics Test; Numeric and Geometric Skills
  ) %>%
  select(id = 1, matches("_16"))
```

```{r, echo = FALSE}
get_descriptives(ncds_16)
plot_cog(ncds_16)
rm(ncds_0016_raw)
```

### Age 50

```{r}
ncds_50_raw <- read_dta("D:/NCDS/50y/ncds_2008_followup.dta")
```

```{r}
ncds_50 <- ncds_50_raw %>%
  mutate(
    fluency_50 = negative_to_na(N8CFANI), # Verbal Fluency (Animal Naming) Test; Verbal (Fluency)
    memory_immediate_50 = negative_to_na(N8CFLISN), # Verbal Learning/ Word List Recall Test (Immediate); Verbal (Memory)     
    memory_delayed_50 = negative_to_na(N8CFLISD), # Verbal Learning/ Word List Recall Test (Delayed); Verbal (Memory) 
    process_speed_50 = negative_to_na(N8CFRC), # Timed Letter Search/ Letter Cancellation Test; Processing Speed
    process_accuracy_50 = negative_to_na(N8CFMIS), # Timed Letter Search/ Letter Cancellation Test; Processing Accuracy (# Mistakes)
  ) %>%
  select(id = 1, matches("_50"))
```

```{r, echo = FALSE}
get_descriptives(ncds_50)
plot_cog(ncds_50)
plot_cor(ncds_50)
```

## Measures of IQ
### Age 7
### Age 11
### Age 16
### Age 50

# British Cohort Study (1970 cohort)
## Individual Cognition Measures
### Age 5
```{r}
bcs_05_raw <- read_dta("D:/BCS70/5y/f699c.dta")
```

```{r}
bcs_05 <- bcs_05_raw %>%
  mutate(
    # f099; Schonell Reading Test; Verbal (Reading); Able to Read (Not cleaned correctly) TO ADD
    reading_05 = negative_to_na(f100), # Schonell Reading Test; Verbal (Reading) (Not cleaned correctly)
    vocabulary_05 = case_when( # English Picture Vocabulary Test; Verbal (Vocabulary)
      f086 == 91 ~ 56 - f087, # Maximum possible correct - errors if no ceiling item
      f085 == 0 ~ 0, # Minimum possible score if no base item (in practice, could be 0-4)
      between(as.numeric(f117), 1, 56) ~ as.numeric(f117), # This is different to that in Moulton et al. (2020)
      TRUE ~ NA_real_
    ), 
    visual_05 = negative_to_na(f119), # Copying Designs Test (CDT); Visual spatial; Note, missingness is less than number who didn't attempt all designs
    drawing_harris_v1_05 = negative_to_na(f113), # Human Figure Drawing; General Ability (Perceptual); 1st figure, Harris Scoring
    drawing_harris_v2_05 = negative_to_na(f114), # Human Figure Drawing; General Ability (Perceptual); 2nd figure, Harris Scoring
    drawing_koppitz_v1_05 = negative_to_na(f115), # Human Figure Drawing; General Ability (Perceptual); 1st figure, Koppitz Scoring
    drawing_koppitz_v2_05 = negative_to_na(f116), # Human Figure Drawing; General Ability (Perceptual); 2nd figure, Koppitz Scoring
    spatial_05 = negative_to_na(f118), # Complete a Profile Test; Spatial Development
  ) %>%
  select(id = 1, matches("_05"))
```

```{r, echo = FALSE}
get_descriptives(bcs_05)
plot_cog(bcs_05)
plot_cor(bcs_05)
rm(bcs_05_raw)
```

### Age 10
```{r}
bcs_10_derived <- read_dta("D:/BCS70/10y/bcs3derived.dta")
bcs_10_raw <- read_dta("D:/BCS70/10y/sn3723.dta")
```

```{r}
bcs_10 <- bcs_10_derived %>%
  mutate(
    reading_10 = negative_to_na(BD3RREAD), # Edinburgh Reading Test; Verbal (Word Recognition)
    maths_10 = negative_to_na(BD3MATHS), # Friendly Maths Test; Mathematics
    maths_10 = negative_to_na(BD3MATHS), 
  ) %>%
  select(id = 1, matches("_10")) %>%
  full_join(get_correct(bcs_10_raw, i8:i97, 
                        comp_sentence_10, 
                        function(x) ifelse(x == 0, 1, 0)), # Pictorial Language Comprehension Test; Verbal; Vocabulary and Sentence Items,
            by = "id") %>%
  full_join(get_correct(bcs_10_raw, i98:i110, 
                        comp_sequence_10), # # Pictorial Language Comprehension Test; Verbal; Sequence Items
            by = "id") %>%
  mutate(comp_total_10 = comp_sentence_10 + comp_sequence_10) %>% # THIS IS GOING TO CAUSE A PROBLEM WHERE ANSWER FOR COMP_SENTENCE_10 BUT NOT COMP_SEQUENCE_10 AND VICE VERSA
  full_join(get_correct(bcs_10_raw, i3815:i3864, spelling_10), # Spelling Dictation Task; Verbal (Spelling)
            by = "id") %>%
  full_join(get_correct(bcs_10_raw, i4201:i4221, bas_sim_10), # BAS Similarities (Words); Verbal (Reasoning); Slightly different sample sizes
            by = "id") %>%
  full_join(get_correct(bcs_10_raw, i3504:i3540, bas_words_10), # BAS Word Definitions; Verbal Knowledge (Acquired and Expressive)
            by = "id") %>%
  full_join(get_correct(bcs_10_raw, i3541:i3574, bas_digits_10), # BAS Word Definitions; Short Term Auditory Memory
            by = "id") %>%
  full_join(get_correct(bcs_10_raw, i3617:i3644, bas_matrices_10), # BAS Matrices; Inductive Non-Verbal Reasoning; Different sample sizes
            by = "id")
```

```{r, echo = FALSE}
get_descriptives(bcs_10)
plot_cog(bcs_10)
plot_cor(bcs_10)
rm(bcs_10_derived, bcs_10_raw)
```


### Age 16
```{r}
bcs_16_reading <- read_dta("D:/BCS70/16y/bcs1986_reading_matrices.dta")
bcs_16_maths <- read_dta("D:/BCS70/16y/bcs70_16-year_arithmetic_data.dta")
bcs_16_raw <- read_dta("D:/BCS70/16y/bcs7016x.dta")
```

```{r}
make_correct <- function(x){
  lbls <- val_labels(x)
  case_when(x == lbls[str_detect(names(lbls), "^correct")] ~ 1,
            x < 0 ~ NA_real_,
            x >= 0 ~ 0)
}

bcs_16 <- bcs_16_reading %>%
  full_join(bcs_16_maths, by = c("BCSID" = "bcsid")) %>%
  mutate(
    ert_skim_16 = negative_to_na(SCR_A), # Edinburgh Reading Test; Verbal (Reading); Skimming
    ert_vocab_16 = negative_to_na(SCR_B), # Edinburgh Reading Test; Verbal (Reading); Skimming
    ert_facts_16 = negative_to_na(SCR_C), # Edinburgh Reading Test; Verbal (Reading); Skimming
    ert_pov_16 = negative_to_na(SCR_D), # Edinburgh Reading Test; Verbal (Reading); Skimming
    ert_comp_16 = negative_to_na(SCR_E), # Edinburgh Reading Test; Verbal (Reading); Skimming
    ert_total_16 = negative_to_na(SCRTOTAL), # Edinburgh Reading Test; Verbal (Reading)
    bas_matrices_16 = negative_to_na(SCR_M), # BAS Matrices; Inductive Non-Verbal Reasoning
    apu_arithmetic_16 = negative_to_na(mathscore), # APU Arithmetic Test; Arithmetic Achievement
  ) %>%
  full_join(bcs_16_raw %>% 
              mutate(across(matches("^cvo"), make_correct)) %>%
              get_correct(cvo1:cvo75, apu_vocab_16), # APU Vocabulary Test; Verbal Vocabulary; Different to Total Score
            by = "id") %>%
  full_join(get_correct(bcs_16_raw, c7a1:c7a100, spelling_a_16), # Spelling Test; Verbal (Spelling); Test A
            by = "id") %>%
  full_join(get_correct(bcs_16_raw, c7b1:c7b100, spelling_b_16), # Spelling Test; Verbal (Spelling); Test B
            by = "id") %>%
  mutate(spelling_total_16 = spell_a_16 + spelling_b_16) %>% # Spelling Test; Verbal (Spelling); Total Score, not same as guide as I clean Tests A and B separately
  select(id = 1, matches("_16"))
```

```{r, echo = FALSE}
get_descriptives(bcs_16)
plot_cog(bcs_16)
plot_cor(bcs_16)
rm(bcs_16_reading, bcs_16_maths, bcs_16_raw)
```


### Age 34
### Age 42
### Age 46

## Measures of IQ
### Age 5
### Age 10
### Age 16
### Age 34
### Age 42
### Age 46

# Millennium Cohort Study (2000/2001 cohort)
## Individual Cognition Measures
### Age 9 Months
### Age 3
### Age 5
### Age 7
### Age 11
### Age 14
### Age 17

## Measures of IQ
### Age 9 Months
### Age 3
### Age 5
### Age 7
### Age 11
### Age 14
### Age 17

# Harmonising Cognitive Measures Across Cohorts